### 2023-04-27 chapter2/chapter2.tex
Fixed typo
Added working draft of last paragraph of introduction

\section{Introduction}
Intrinsically disordered regions (IDRs) are segments of proteins which lack stable three-dimensional structures and instead exist as ensembles of rapidly interconverting conformations. As a result of this structural hetereogeneity, IDRs can interact with diverse binding partners. Often these interactions have high specificity but moderate affinity, which permits the efficient propagation of signals by rapid binding and dissociation~\cite{Zhou2012, Wright2014}. Furthermore, as IDRs readily expose their polypeptide chains, they are enriched in recognition motifs for post-translational modifications which allow environmental or physiological conditions to modulate their interactions. Accordingly, IDRs often act as the ``hubs'' of complex signaling networks by integrating signals from diverse pathways and coordinating interactions~\cite{Dunker2005, Buljan2012}. However, as IDRs are pervasive in eukaryotic proteomes, with estimates of the fractions of disordered residues in the human, mouse, and fruit fly proteomes ranging between 22 and 24\%~\cite{Piovesan2020, Piovesan2022}, they are involved in diverse processes~\cite{vanderLee2014} including transcriptional regulation~\cite{Liu2006} and the formation of biomolecular condensates~\cite{Banani2017}.

@@

% IV. Thus, in this work Explicit phylogenetic models of trait evolution can reveal patterns of evolution with fewer assumptions
%    A. Previous analyses compared observed distributions of features in observed IDRs to the distributions generated by simulated counterparts
%    B. Differences were interpreted as evidence of selection on features
%        1. However, evidence of deviation from the null model rather than evidence of constraint necessarily
%        2. Null distribution generated by a model of evolution where the composition of sequences stochastically evolve towards a specific equilibrium distribution
%        3. Previous studies have identified IDRs with varying compositional preferences, so a single equilibrium distribution is unlikely to globally describe observed patterns of sequence evolution in IDRs
%    C. Thus in this work we explore use of various statistical approaches that directly measure rates of change and constraint of molecular features indepedent of specific model of sequence evolution
%        1. Place these evolutionary analyses in a statistical phylogenetic framework through the application of the Brownian motion and Ornstein-Uhlenbeck models of trait evolution.
%        2. Though these models have own assumptions, they are macroscopic descriptions of continuous trait evolution and therefore independent of any assumptions about evolution at the level sequences
%        3. Still have their own assumptions of course, but more general and robust to deviations
%        4. We view these approaches as complementary, each having different trade-offs between specificity and generality of testing specific hypotheses of disorder evolution
%        5. We also explore hybrid approaches where the evolutionary rates obtained from phylogenetic models are used as a statistic
%
% V. Therefore, in this work we apply evolutionary models of sequence and trait evolution to the sequences of IDRs and their of associated molecular feature to dissect the forces driving their evolution. We find IDRs exhibit unique patterns of evolution at the level of sequences. We find intrinsic disorder is a dynamically evolving property in some regions, and changes in disorder preference are correlated with changes in the complexity. We find that IDRs are unconstrained along several axis of trait evolution, but they also display

>>>>

\section{Introduction}
Intrinsically disordered regions (IDRs) are segments of proteins which lack stable three-dimensional structures and instead exist as ensembles of rapidly interconverting conformations. As a result of this structural heterogeneity, IDRs can interact with diverse binding partners. Often these interactions have high specificity but moderate affinity, which permits the efficient propagation of signals by rapid binding and dissociation~\cite{Zhou2012, Wright2014}. Furthermore, as IDRs readily expose their polypeptide chains, they are enriched in recognition motifs for post-translational modifications which allow environmental or physiological conditions to modulate their interactions. Accordingly, IDRs often act as the ``hubs'' of complex signaling networks by integrating signals from diverse pathways and coordinating interactions~\cite{Dunker2005, Buljan2012}. However, as IDRs are pervasive in eukaryotic proteomes, with estimates of the fractions of disordered residues in the human, mouse, and fruit fly proteomes ranging between 22 and 24\%~\cite{Piovesan2020, Piovesan2022}, they are involved in diverse processes~\cite{vanderLee2014} including transcriptional regulation~\cite{Liu2006} and the formation of biomolecular condensates~\cite{Banani2017}.

@@

V. Therefore, in this work we apply a series of phylogenetic models to the sequences and molecular features of a set of orthologous IDRs identified in the \textit{Drosophila} genome to dissect the forces driving their evolution. Though the previous approach relied on simulations to generate the null distribution for a hypothesis of no constraint, our analyses of their molecular features instead leverage a fully statistical phylogenetic comparative framework. By comparing models of constrained and unconstrained continuous trait evolution, \textit{i.e.} the Brownian motion and Ornstein-Uhlenbeck models, respectively, we can demonstrate evidence of selective constraint independent of any assumptions about the underlying process of sequence evolution. However, we also explore hybrid approaches that combine simulations with phylogenetic comparative methods to test increasingly refined models of IDR evolution. Our results show IDRs exhibit unique patterns of amino acid substitution where structure-promoting residues are under less constraint in IDRs despite their relative scarcity. Additionally, by treating intrinsic disorder as trait a global analyses of the disorder as trait in some proteins disorder is a dynamically evolving property, with changes in predicted disorder strongly correlated with changes in many distributed features, including hydrophobicity and sequence complexity. Furthermore, we find that while IDRs are unconstrained along several axes of trait evolution, they also reveal conserved patterns of feature constraint which are associated with function, which suggests the previously
